---
title: "FoldX Error Model: Organize data into dataframes"
author: "Yesol Sapozhnikov"
date: "1/5/2022"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


FoldX datasets from MD snapshots are in /data directory. These text files are read and organized into arrays for further manipulation and analysis.

Input data location: 
MD+FoldX datasets: data/fold_data/, data/bind_data/
FoldX-only datasets: data/fold_xstal_data/, data/bind_xstal_data/
*experimental ddg values in the original dataset files were incorrect. Corrected numbers are in data/binding_exp/
DSSP outputs: data/dssp/

Below chunk reads in the data text files and creates 
```{r}
path_f <- "../data/fold_data/"
path_b <- "../data/bind_data/"
dns_f <- sort(list.files(path_f))
dns_b <- sort(list.files(path_b))
fns_f <- list.files(paste(path_f, dns_f[1], sep = "")) #files in each sys dir are identical, so grab them from 1bni
fns_b <- list.files(paste(path_b, dns_b[1], sep = ""))
sys_names_f <- sapply(strsplit(dns_f, "_"), "[", 1)
sys_names_b <- sapply(strsplit(dns_b, "_"), "[", 1)


######## FOLD DATA ########

# Read in ddg_exp.txt for sys names, muts, exp ddg values
list_sys <- list()
list_mut <- list()
list_ddg_exp <- list()
for (i in 1:length(dns_f)){
  list_sys[[i]] <- read.table(paste(path_f, dns_f[i], "/ddg_exp.txt", sep = ""))[1]
  list_mut[[i]] <- read.table(paste(path_f, dns_f[i], "/ddg_exp.txt", sep = ""))[2]
  list_ddg_exp[[i]] <- read.table(paste(path_f, dns_f[i], "/ddg_exp.txt", sep = ""))[3]
}
sys <- tolower(unlist(list_sys))
mut <- unlist(list_mut)
ddg_exp <- unlist(list_ddg_exp)
list_exp <- list("systems" = sys, "mut" = mut, "ddg_exp" = ddg_exp)
fold <- as.data.frame(list_exp)
# Get rest of the energy terms into the df
col_names_f <- fns_f
col_names_f <- sub("[.]txt", "", col_names_f)
#col_names_f <- col_names_f[!(col_names_f %in% c("ddg_exp","mut","total"))] #If want to exclude total
col_names_f <- col_names_f[!(col_names_f %in% c("ddg_exp","mut"))] #include total
list_vals <- list()
for (i in 1:length(col_names_f)){
  vals_i <- list()
  for (j in 1:length(dns_f)){
    vals_i[[j]] <- read.table(paste(path_f,dns_f[j],"/",col_names_f[i],".txt",sep=""))[1]
  }
  list_vals[[col_names_f[i]]] <- unlist(vals_i)
}
# Combine into a single df
fold <- cbind(fold, as.data.frame(list_vals))
# Add "error" columne
fold$error <- abs(fold$total-fold$ddg_exp)


######## BIND DATA #########

# Repeating the code with slight modification due to the slight different file str
# e.g. ddg_exp.txt does not contain sys, mut info as in fold data
systems <- list()
for (i in 1:length(sys_names_b)){
  systems <- append(systems,
                    rep(sys_names_b[i],nrow(read.table(paste(path_b,"/",dns_b[i],"/mut.txt",sep="")))))
}
list_sys <- tolower(unlist(systems))
# Do mut and ddg_exp for begibnning of the df:
list_mut <- list()
list_ddg_exp <- list()
for (i in 1:length(dns_b)){
  list_mut[[i]] <- read.table(paste(path_b,dns_b[i],"/mut.txt",sep=""))
}

# Numbers in ddg_exp.txt had to be corrected. Received as separate files. Read them in and add to the dataframe:
path_exp_b <- "../data/binding_exp/"
for (i in 1:length(dns_b)){
  list_ddg_exp[[i]] <- read.table(paste(path_exp_b,sys_names_b[i],"/",sys_names_b[i],"_ddg_exp.txt",
                                        sep = ""))}


mut <- unlist(list_mut)
ddg_exp <- unlist(list_ddg_exp)
list_exp <- list("systems"=list_sys,"mut"=mut,"ddg_exp"=ddg_exp)
bind <- as.data.frame(list_exp)

# Get rest of the values into the df
# first columns - values
col_names_b <- fns_b
col_names_b <- sub("[.]txt","",col_names_b)
#col_names_b <- col_names_b[!(col_names_b %in% c("ddg_exp","mut","total"))] #If want to exclude total
col_names_b <- col_names_b[!(col_names_b %in% c("ddg_exp","mut"))] #include total
list_vals <- list()
for (i in 1:length(col_names_b)){
  vals_i <- list()
  for (j in 1:length(dns_b)){
    vals_i[[j]] <- read.table(paste(path_b,dns_b[j],"/",col_names_b[i],".txt",sep=""))[1]
  }
  list_vals[[col_names_b[i]]] <- unlist(vals_i)
}
# Combine into a single df
bind <- cbind(bind, as.data.frame(list_vals))
# Add "error" columne
bind$error <- abs(bind$total-bind$ddg_exp)


# Save sys, mut, ddg_exp columns for FoldX-only dataframes later:
foldx <- fold[c("systems","mut","ddg_exp")]
bindx <- bind[c("systems","mut","ddg_exp")]


##### MD snapshots SD values #######

# sd (second col of the data files) - Fold
list_sd_f <- list()
for (i in 1:length(col_names_f)){
  sd_i <- list()
  for (j in 1:length(dns_f)){
    sd_i[[j]] <- read.table(paste(path_f, dns_f[j],"/",col_names_f[i],".txt",sep=""))[2]
  }
  list_sd_f[[paste(col_names_f[i],"_sd",sep="")]] <- unlist(sd_i)
}
fold <- cbind(fold, as.data.frame(list_sd_f))

# Bind
list_sd_b <- list()
for (i in 1:length(col_names_b)){
  sd_i <- list()
  for (j in 1:length(dns_b)){
    sd_i[[j]] <- read.table(paste(path_b, dns_b[j],"/",col_names_b[i],".txt",sep=""))[2]
  }
  list_sd_b[[paste(col_names_b[i],"_sd",sep="")]] <- unlist(sd_i)
}
bind <- cbind(bind, as.data.frame(list_sd_b))


# Remove energy terms that are all zero
zeroes_f <- vector()
for (i in 3:length(colnames(fold))){
  if (sum(fold[,i])==0){
    zeroes_f <- c(zeroes_f, i)
  }
}
fold <- fold[-zeroes_f]

zeroes_b <- vector()
for (i in 3:length(colnames(bind))){
  if (sum(bind[,i])==0){
    zeroes_b <- c(zeroes_b, i)
  }
}
bind <- bind[-zeroes_b]


###### Fill in FoldX-only data ######

path_fx <- "../data/fold_xstal_data/"
path_bx <- "../data/bind_xstal_data/"
dns_fx <- sort(list.files(path_fx))
dns_bx <- sort(list.files(path_bx))
fns_fx <- list.files(paste(path_fx, dns_fx[1], sep=""))
fns_bx <- list.files(paste(path_bx, dns_bx[1], sep=""))
sys_names_fx <- sapply(strsplit(dns_f,"_"),"[",1)
sys_names_bx <- sapply(strsplit(dns_bx,"_"),"[",1)

###### FOLD DATA ######
# sys, mut, ddg_exp info saved as foldx, bindx dataframes.
# read in foldX output values
col_names_fx <- fns_fx
col_names_fx <- sub("ddg_0_","",col_names_fx)
col_names_fx <- sub("[.]dat","",col_names_fx)
list_vals <- list()
for (i in 1:length(col_names_fx)){
  vals_i <- list()
  for (j in 1:length(dns_fx)){
    vals_i[[j]] <- read.table(paste(path_fx,dns_fx[j],"/",fns_fx[i],sep=""))[1]
  }
  list_vals[[col_names_fx[i]]] <- unlist(vals_i)
}
df_vals_fx <- as.data.frame(list_vals) # convert list of values as df for cleanup
# rows 676-702 are duplicates of 27 rows of 2abd. Delete these rows.
df_vals_fx <- df_vals_fx[1:675,]
# Combine into a single df
foldx <- cbind(foldx, df_vals_fx)
# Add "error" columne
foldx$error <- abs(foldx$total-foldx$ddg_exp)

# Remove energy terms that have only zeroes:
zeroes_fx <- vector()
for (i in 3:length(colnames(foldx))){
  if (sum(foldx[,i])==0){
    zeroes_fx <- c(zeroes_fx, i)
  }
}
foldx <- foldx[-zeroes_fx]

####### BIND DATA ######
col_names_bx <- fns_bx
col_names_bx <- sub("ddg_000_","",col_names_bx)
col_names_bx <- sub("[.]dat","",col_names_bx)
list_vals <- list()
for (i in 1:length(col_names_bx)){
  vals_i <- list()
  for (j in 1:length(dns_bx)){
    vals_i[[j]] <- read.table(paste(path_bx,dns_bx[j],"/",fns_bx[i],sep=""))[1]
  }
  list_vals[[col_names_bx[i]]] <- unlist(vals_i)
}
# Combine into a single df
bindx <- cbind(bindx, as.data.frame(list_vals))
# Add "error" columne
bindx$error <- abs(bindx$total-bindx$ddg_exp)

# Remove energy terms that have only zeroes:
zeroes_bx <- vector()
for (i in 3:length(colnames(bindx))){
  if (sum(bindx[,i])==0){
    zeroes_bx <- c(zeroes_bx, i)
  }
}
bindx <- bindx[-zeroes_bx]

# Remove columns that will not be in the model search:
omit_fx <- which(names(foldx) %in% c("ddg_exp","systems","mut"))
omit_bx <- which(names(bindx) %in% c("ddg_exp","systems","mut"))
foldx_tr <- foldx[-omit_fx]
bindx_tr <- bindx[-omit_bx]
```



## Calculate and add other amino acid characteristics

Put reference values for VDW volumes, pK, hydrophobicity index, and charge in a named list. This will be looked up in calculating changes of these values in each mutation in dataset.
```{r}
# AA and their properties (ref: textbook, Sigma Aldrich reference chart)
AAs <- c("A","R","N","D","C","Q","E","G","H","I","L","K","M","F","P","S","T","W","Y","V")
# vdw volume in angstrom^3
vdw <- c(88.6,173.4,114.1,111.1,108.5,143.8,138.4,60.1,153.2,166.7,166.7,168.6,162.9,189.9,112.7,89.0,116.1,227.8,193.6,140.0)
pK <- c(NA,12.48,NA,3.65,8.18,NA,4.25,NA,6,NA,NA,10.53,NA,NA,NA,NA,NA,NA,NA,NA) # pK of charged groups
HI <- c(41,-14,-28,-55,49,-10,-31,0,8,99,97,-23,74,100,-46,-5,13,97,63,76) # Pro: at pH2 
charge <- c(0,1,0,-1,0,0,-1,0,1,0,0,1,0,0,0,0,0,0,0,0) # plain electrical charge
listAA <- list(vdw, pK, HI, charge)
for (i in 1:length(listAA)){
  names(listAA[[i]]) <- AAs
}
names(listAA) <- c("vdw", "pK", "HI", "charge")

# Create new columns for AA change information for characteristics such as changes in VDW volumne, charge, etc.
# Separate out AA changes:
fold$AAfrom <- sapply(as.character(fold$mut), function(x) strsplit(x, "")[[1]][1])
fold$AAto <- sapply(as.character(fold$mut), function(x) strsplit(x, "")[[1]][length(strsplit(x, "")[[1]])])
fold$chain <- sapply(as.character(fold$mut), function(x) strsplit(x, "")[[1]][2])
fold$residue <- gsub('\\D', '', fold$mut)
fold$ch_res <- paste(fold$chain, fold$residue, sep = "") # for matching with dssp rows later

bind$AAfrom <- sapply(as.character(bind$mut), function(x) strsplit(x, "")[[1]][1])
bind$AAto <- sapply(as.character(bind$mut), function(x) strsplit(x, "")[[1]][length(strsplit(x, "")[[1]])])
bind$chain <- sapply(as.character(bind$mut), function(x) strsplit(x, "")[[1]][2])
bind$residue <- gsub('\\D', '', bind$mut)
bind$ch_res <- paste(bind$chain, bind$residue, sep = "")


# Mutations to or from proline
fold$P <- ifelse(fold$AAfrom == "P" & fold$AAto != "P", 1, 
                     ifelse(fold$AAfrom != "P" & fold$AAto == "P", 1, 0))
bind$P <- ifelse(bind$AAfrom == "P" & bind$AAto != "P", 1, 
                     ifelse(bind$AAfrom != "P" & bind$AAto == "P", 1, 0))
# Size change (using vdw volume)
fold$volchange <- abs(listAA$vdw[fold$AAto] - listAA$vdw[fold$AAfrom])
bind$volchange <- abs(listAA$vdw[bind$AAto] - listAA$vdw[bind$AAfrom])
# Hydrophobicity change
fold$HIchange <- abs(listAA$HI[fold$AAto] - listAA$HI[fold$AAfrom])
bind$HIchange <- abs(listAA$HI[bind$AAto] - listAA$HI[bind$AAfrom])
# Charge change
fold$charge <- abs(listAA$charge[fold$AAto] - listAA$charge[fold$AAfrom])
bind$charge <- abs(listAA$charge[bind$AAto] - listAA$charge[bind$AAfrom])
```



Secondary structure info was obtained by running DSSP (DSSP usage: https://swift.cmbi.umcn.nl/gv/dssp/DSSP_3.html). 

For Relative Solven Accessibility (RSA) calculation, we take ASA values from DSSP output and use methods described in: Tien MZ, Meyer AG, Sydykova DK et al. Maximum Allowed Solvent Accessibilites of Residues in Proteins. PLOS ONE 2013;8:e80635

DSSP output files have been modified in Excel to cut header and separate out columns before saving as CSV files that will be read in below.

```{r}
# FOLD
# Read in each dssp files into a list of dataframes
dssps_f <- list()
for (i in 1:length(sys_names_f)) {
  dssps_f[[i]] <- read.csv(paste("../data/dssp/fold/sys",
                                 sys_names_f[i], ".dssp.csv", sep = ""))
  dssps_f[[i]]$ch_res <- paste(dssps_f[[i]]$CHAIN, dssps_f[[i]]$RESIDUE, sep = "") 
  dssps_f[[i]]$STR <- sapply(dssps_f[[i]]$STR, as.character) # without it, reads as factor
}
names(dssps_f) <- sys_names_f
# Match each mut to DSSP entry and add ACC and sec str info to the master df
for (r in 1:nrow(fold)){
  dssp_tmp <- dssps_f[[fold$systems[r]]]
  chres_tmp <- fold$ch_res[r]
  if (chres_tmp %in% dssp_tmp$ch_res){
    fold$str[r] <- dssp_tmp[which(dssp_tmp$ch_res==chres_tmp),]$STR
    fold$ACC[r] <- dssp_tmp[which(dssp_tmp$ch_res==chres_tmp),]$ACC
  } else {
    print(paste("(fold) no match for:", fold$systems[r], fold$ch_res[r], 
                "- Will remove row", r, sep = " "))
    fold$str[r] <- NA
    fold$ACC[r] <- NA
  }
}
# Remove "no match" entries from master df (gets error later on otherwise) (there are only 3)
fold <- na.omit(fold)

# BIND
# Read in each dssp files into a list of data frames
dssps_b <- list()
for (i in 1:length(sys_names_b)) {
  dssps_b[[i]] <- read.csv(paste("../data/dssp/bind/sys",
                                 sys_names_b[i], ".dssp.csv", sep = ""))
  dssps_b[[i]]$ch_res <- paste(dssps_b[[i]]$CHAIN, dssps_b[[i]]$RESIDUE, sep = "")
  dssps_b[[i]]$STR <- sapply(dssps_b[[i]]$STR, as.character)
}
names(dssps_b) <- sys_names_b
# Match each mut to DSSP entry and add ACC and sec str info to the master df
for (r in 1:nrow(bind)){
  dssp_tmp <- dssps_b[[bind$systems[r]]]
  chres_tmp <- bind$ch_res[r]
  if (chres_tmp %in% dssp_tmp$ch_res){
    bind$str[r] <- dssp_tmp[which(dssp_tmp$ch_res==chres_tmp),]$STR
    bind$ACC[r] <- dssp_tmp[which(dssp_tmp$ch_res==chres_tmp),]$ACC
  } else {
    print(paste("(bind) no match for:", bind$systems[r], bind$ch_res[r],
                "- Will remove row", r, sep = " "))
    bind$str[r] <- NA
    bind$ACC[r] <- NA
  }
}
bind <- na.omit(bind)

# RSA calculation
# Table S1 values (Tien et al)
# Recommended (ALLOWED category) maximum SA values - do empirical values
max_emp <- c(121,265,187,187,148,214,214,97,216,195,191,230,203,228,154,143,163,264,255,166)
names(max_emp) <- AAs
fold$RSA <- fold$ACC/max_emp[fold$AAfrom]
bind$RSA <- bind$ACC/max_emp[bind$AAfrom]
fold$str[fold$str==""] <- "NONE"
bind$str[bind$str==""] <- "NONE"
fold$str <- as.factor(fold$str)
bind$str <- as.factor(bind$str)
```

Remove columns that will not be in the model search:
```{r}
omit_f <- which(names(fold) %in% c("ddg_exp","systems","mut","AAfrom","AAto","chain","residue","ch_res","ACC"))
omit_b <- which(names(bind) %in% c("ddg_exp","systems","mut","AAfrom","AAto","chain","residue","ch_res","ACC"))
fold_tr <- fold[-omit_f]
bind_tr <- bind[-omit_b]
```





After running above codes, dataframes we get are:
- foldx/bindx: FoldX-only values with system, mut, ddg_exp info
- foldx_tr/bindx_tr: systems, mut, ddg_exp removed from above. Use for model training
- fold/bind: FoldX values + sd (from MD sim) + other info (RSA, structure, etc)
- fold_tr/bind_tr: Use for model training


Save dataframes in outputs/
```{r}
write.csv(foldx, "../outputs/table_fxonly_f.csv", row.names = F)
write.csv(bindx, "../outputs/table_fxonly_b.csv", row.names = F)
write.csv(foldx_tr, "../outputs/tr_table_fxonly_f.csv", row.names = F)
write.csv(bindx_tr, "../outputs/tr_table_fxonly_b.csv", row.names = F)

write.csv(fold, "../outputs/table_full_f.csv", row.names = F)
write.csv(bind, "../outputs/table_full_b.csv", row.names = F)
write.csv(fold_tr, "../outputs/tr_table_full_f.csv", row.names = F)
write.csv(bind_tr, "../outputs/tr_table_full_b.csv", row.names = F)
```






