---
title: "Model Summary"
author: "Yesol Sapozhnikov"
date: '2022-06-14'
output: html_document
---

<style type="text/css">
  body{
  font-size: 10pt;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, include = FALSE}
# load packages
library(tidyverse)

# Full tables:
fold <- read.csv("../outputs/table_full_f.csv")
bind <- read.csv("../outputs/table_full_b.csv")


# FoldX-only tables:
foldx <- read.csv("../outputs/table_fxonly_f.csv")
bindx <- read.csv("../outputs/table_fxonly_b.csv")

foldx <- foldx[-c(449, 642, 643),]

# selected models from stepwise & best subset selection
best_mods <- readRDS("../outputs/best_mods_all.rds")

# source LOsysOCV function
source("../scripts/functions.R")
```


## Model performance comparison

Fold dataset
```{r, warning=FALSE}
# Must create separate datasets for mod2 that contains fx-only data plus biochem params.
mod2_data_f <- fold %>% 
  select(P, volchange, HIchange, charge, str, RSA) %>%
  bind_cols(., foldx)
mod2_data_b <- bind %>%
  select(P, volchange, HIchange, charge, str, RSA) %>%
  bind_cols(., bindx)
datasets_ls <- list(foldx, mod2_data_f, fold, fold, fold, bindx, mod2_data_b, bind, bind, bind)
sum_table <- data.frame(models = str_split(names(best_mods[1:5]), "_", simplify = T)[,1])
for (i in 1:5){
  sum_table$adj_r2[i] <- summary(best_mods[[i]])$adj.r.squared
  out <- loocv(formula(best_mods[[i]]), datasets_ls[[i]])
  sum_table$coverage[i] <- out$coverage
  sum_table$med_upr_bd[i] <- out$med_upr_bd
}
sum_table
```

Bind dataset
```{r, warning = FALSE}
sum_table <- data.frame(models = str_split(names(best_mods[6:10]), "_", simplify = T)[,1])
for (i in 6:10){
  sum_table$adj_r2[i-5] <- summary(best_mods[[i]])$adj.r.squared
  out <- loocv(formula(best_mods[[i]]), datasets_ls[[i]])
  sum_table$coverage[i-5] <- out$coverage
  sum_table$med_upr_bd[i-5] <- out$med_upr_bd
}
sum_table
```

## Model details: Fold

### Model 5 (MD+FX dataset with all predictors)

```{r}
summary(best_mods$mod5_fold)
```

### Model 4 (MD+FX dataset with energy terms and biochemical properties)

```{r}
summary(best_mods$mod4_fold)
```

### Model 3 (MD+FX dataset with energy terms only)

```{r}
summary(best_mods$mod3_fold)
```

### Model 2 (FX dataset with energy terms and biochemical properties)

```{r}
summary(best_mods$mod2_fold)
```

### Model 1 (FX dataset with energy terms only)

```{r}
summary(best_mods$mod1_fold)
```


## Model details: Bind

### Model 5 (MD+FX dataset with all predictors)

```{r}
summary(best_mods$mod5_bind)
```

### Model 4 (MD+FX dataset with energy terms and biochemical properties)

```{r}
summary(best_mods$mod4_bind)
```

### Model 3 (MD+FX dataset with energy terms only)

```{r}
summary(best_mods$mod3_bind)
```

### Model 2 (FX dataset with energy terms and biochemical properties)

```{r}
summary(best_mods$mod2_bind)
```

### Model 1 (FX dataset with energy terms only)

```{r}
summary(best_mods$mod1_bind)
```

# Compare BIC

```{r}
bics <- sapply(best_mods, BIC)
bic_df <- data.frame(model = str_sub(names(bics), 1, 4),
                     bic = round(bics),
                     energy = str_sub(names(bics), 6, 9))
bic_df %>%
  ggplot(aes(x = model, y = bic, color = energy, group = energy)) +
  geom_point() +
  geom_line() +
  geom_text(aes(label = bic), vjust = 2, color = "black") +
  scale_color_manual(values = c("red", "steelblue")) +
  theme_bw()
ggsave("../figures/BIC_comparison.png")
```




