---
title: "Application"
author: "Yesol Sapozhnikov"
date: "2023-04-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

Use this notebook to obtain uncertainty associated with FoldX-calculated ddG, i.e. the upper bound of the prediction interval around predicted error (see manuscript for details).

How to use:
1. Choose the appropriate model
2. Edit the input dataframe: replace the values for each column. Do not change the column names as they are the coefficient names in the lm object representing the model. Each row represents values associated with each mutation.
3. Run the chunk which will output the result.

Result interpretation:
`out` is an n x 3 numeric array output by `predict()` function. "fit" is the predicted error, while "lwr" and "upr" are the lower and upper bounds of the prediction interval around the point estimate error. Because of how we defined error, the lower bound is not meaningful. 
`upr_bound` is a numeric vector, showing just the uppoer bounds of the prediction interval. +/- upr_bound values capture 95% of actual error. Or put another way, FoldX ddG +/- upr_bound would contain the true (experimental) ddG (see figure 2 in the manuscript for graphical illustration).


## Model 1 for folding energy

Use this model for ddG calculated from a single experimental structure. Expected input data are ddG values for each of these FoldX terms: energy_vdw, energy_vdwclash, entropy_sidec, sideHbond, total.

```{r}
# Create the input data. Replace the simulated values with your own data
dat <- data.frame(
  energy_vdw = rnorm(3, 1, 1),
  energy_vdwclash = rnorm(3, 1, 1),
  entropy_sidec = rnorm(3, 1, 1),
  sideHbond = rnorm(3, 1, 1),
  total = rnorm(3, 1, 1)
)
# Read in the model object
mod1_fold <- readRDS("../outputs/best_mods_all.rds")$mod1_fold
# Predict and output the result
out <- predict(mod1_fold, newdata = dat, interval = "prediction")
upr_bound <- out[,"upr"]
print(out)
cat("upper bounds: ", upr_bound)
```


## Model 1 for binding energy

Use this model for ddG calculated from a single experimental structure. Expected input data are ddG values for each of these FoldX terms: energy_ionisation, energy_solvh, energy_vdw, energy_vdwclash, entropy_sidec.
```{r}
# Create the input data. Replace the simulated values with your own data
dat <- data.frame(
  energy_ionisation = rnorm(3, 1, 1),
  energy_solvh = rnorm(3, 1, 1),
  energy_vdw = rnorm(3, 1, 1),
  energy_vdwclash = rnorm(3, 1, 1),
  entropy_sidec = rnorm(3, 1, 1)
)
# Read in the model object
mod1_bind <- readRDS("../outputs/best_mods_all.rds")$mod1_bind
# Predict and output the result
out <- predict(mod1_bind, newdata = dat, interval = "prediction")
upr_bound <- out[,"upr"]
print(out)
cat("upper bounds: ", upr_bound)
```


## Model 2 for folding energy

Use this model for ddG calculated from a single experimental structure. 
Expected input data are:
- ddG values for each of these FoldX terms: energy_vdw, energy_vdwclash, entropy_sidec, sideHbond, total
- Mutation to/from proline (0 for false, 1 for true)
- Secondary structure according to DSSP ("B", "E", "G", "H", "NONE", "S", "T")
- RSA calcuated as described in the manuscript and 01_generate-organized-tables.Rmd, using a method from Tien et al.

```{r}
# Create the input data. Replace the simulated values with your own data
dat <- data.frame(
  energy_vdw = rnorm(3, 1, 1),
  energy_vdwclash = rnorm(3, 1, 1),
  entropy_sidec = rnorm(3, 1, 1),
  sideHbond = rnorm(3, 1, 1),
  total = rnorm(3, 1, 1),
  P = c(0, 0, 1),
  str = c("B", "E", "H"),
  RSA = c(.2, .2, .8)
)
# Read in the model object
mod2_fold <- readRDS("../outputs/best_mods_all.rds")$mod2_fold
# Predict and output the result
out <- predict(mod2_fold, newdata = dat, interval = "prediction")
upr_bound <- out[,"upr"]
print(out)
cat("upper bounds: ", upr_bound)
```


## Model 2 for binding energy

Use this model for ddG calculated from a single experimental structure. 
Expected input data are:
- ddG values for each of these FoldX terms: electro, energy_solvh, energy_vdw, energy_vdwclash, entropy_sidec
- Mutation to/from proline (0 for false, 1 for true)
- Secondary structure according to DSSP ("B", "E", "G", "H", "NONE", "S", "T")
- RSA calcuated as described in the manuscript and 01_generate-organized-tables.Rmd, using a method from Tien et al.
- charge changes (absolute value of the difference between the charge of wildtype AA and the charge of the mutated AA, can be 0, 1, or 2)

```{r}
# Create the input data. Replace the simulated values with your own data
dat <- data.frame(
  electro = rnorm(3, 1, 1),
  energy_solvh = rnorm(3, 1, 1),
  energy_vdw = rnorm(3, 1, 1),
  energy_vdwclash = rnorm(3, 1, 1),
  entropy_sidec = rnorm(3, 1, 1),
  P = c(0, 0, 1),
  str = c("B", "E", "H"),
  RSA = c(.2, .2, .8),
  charge = c(0, 1, 2)
)
# Read in the model object
mod2_bind <- readRDS("../outputs/best_mods_all.rds")$mod2_bind
# Predict and output the result
out <- predict(mod2_bind, newdata = dat, interval = "prediction")
upr_bound <- out[,"upr"]
print(out)
cat("upper bounds: ", upr_bound)
```


## Model 3 for folding energy

Use this model for average ddG from MD + FoldX workflow.
Expected input data are averaged ddG values for each of these FoldX terms: energy_solvh, energy_vdw, energy_vdwclash, entropy_mainc, entropy_sidec, total

```{r}
# Create the input data. Replace the simulated values with your own data
dat <- data.frame(
  energy_solvh = rnorm(3, 1, 1),
  energy_vdw = rnorm(3, 1, 1),
  energy_vdwclash = rnorm(3, 1, 1),
  entropy_mainc = rnorm(3, 1, 1),
  entropy_sidec = rnorm(3, 1, 1),
  total = rnorm(3, 1, 1)
)
# Read in the model object
mod3_fold <- readRDS("../outputs/best_mods_all.rds")$mod3_fold
# Predict and output the result
out <- predict(mod3_fold, newdata = dat, interval = "prediction")
upr_bound <- out[,"upr"]
print(out)
cat("upper bounds: ", upr_bound)
```


## Model 3 for binding energy

Use this model for average ddG from MD + FoldX workflow.
Expected input data are averaged ddG values for each of these FoldX terms: energy_vdw, energy_vdwclash, entropy_sidec, sideHbond.

```{r}
# Create the input data. Replace the simulated values with your own data
dat <- data.frame(
  energy_vdw = rnorm(3, 1, 1),
  energy_vdwclash = rnorm(3, 1, 1),
  entropy_sidec = rnorm(3, 1, 1),
  sideHbond = rnorm(3, 1, 1)
)
# Read in the model object
mod3_bind <- readRDS("../outputs/best_mods_all.rds")$mod3_bind
# Predict and output the result
out <- predict(mod3_bind, newdata = dat, interval = "prediction")
upr_bound <- out[,"upr"]
print(out)
cat("upper bounds: ", upr_bound)
```


## Model 4 for folding energy

Use this model for average ddG from MD + FoldX workflow.
Expected input data are:
- Averaged ddG values for each of these FoldX terms: energy_solvh, energy_vdw, energy_vdwclash, entropy_mainc, entropy_sidec, total
- Mutation to/from proline (0 for false, 1 for true)
- Secondary structure according to DSSP ("B", "E", "G", "H", "NONE", "S", "T")

```{r}
# Create the input data. Replace the simulated values with your own data
dat <- data.frame(
  energy_solvh = rnorm(3, 1, 1),
  energy_vdw = rnorm(3, 1, 1),
  energy_vdwclash = rnorm(3, 1, 1),
  entropy_mainc = rnorm(3, 1, 1),
  entropy_sidec = rnorm(3, 1, 1),
  total = rnorm(3, 1, 1),
  P = c(0, 0, 1),
  str = c("B", "E", "H")
)
# Read in the model object
mod4_fold <- readRDS("../outputs/best_mods_all.rds")$mod4_fold
# Predict and output the result
out <- predict(mod4_fold, newdata = dat, interval = "prediction")
upr_bound <- out[,"upr"]
print(out)
cat("upper bounds: ", upr_bound)
```


## Model 4 for binding energy

Use this model for average ddG from MD + FoldX workflow.
Expected input data are:
- Averaged ddG values for each of these FoldX terms: energy_vdw, energy_vdwclash, entropy_sidec
- Mutation to/from proline (0 for false, 1 for true)
- Secondary structure according to DSSP ("B", "E", "G", "H", "NONE", "S", "T")
- RSA calcuated as described in the manuscript and 01_generate-organized-tables.Rmd, using a method from Tien et al.
- charge changes (absolute value of the difference between the charge of wildtype AA and the charge of the mutated AA, can be 0, 1, or 2)

```{r}
# Create the input data. Replace the simulated values with your own data
dat <- data.frame(
  energy_vdw = rnorm(3, 1, 1),
  energy_vdwclash = rnorm(3, 1, 1),
  entropy_sidec = rnorm(3, 1, 1),
  P = c(0, 0, 1),
  str = c("B", "E", "H"),
  RSA = c(.2, .2, .8),
  charge = c(0, 1, 2)
)
# Read in the model object
mod4_bind <- readRDS("../outputs/best_mods_all.rds")$mod4_bind
# Predict and output the result
out <- predict(mod4_bind, newdata = dat, interval = "prediction")
upr_bound <- out[,"upr"]
print(out)
cat("upper bounds: ", upr_bound)
```


## Model 5 for folding energy

Use this model for average ddG from MD + FoldX workflow.
Expected input data are:
- Averaged ddG values for each of these FoldX terms: energy_vdw, energy_vdwclash, entropy_sidec
- Standard deviation of the averaged ddG for total (named "total_sd")
- Mutation to/from proline (0 for false, 1 for true)
- Secondary structure according to DSSP ("B", "E", "G", "H", "NONE", "S", "T")

```{r}
# Create the input data. Replace the simulated values with your own data
dat <- data.frame(
  energy_vdw = rnorm(3, 1, 1),
  energy_vdwclash = rnorm(3, 1, 1),
  entropy_mainc = rnorm(3, 1, 1),
  entropy_sidec = rnorm(3, 1, 1),
  total_sd = rnorm(3, 0, 1),
  P = c(0, 0, 1),
  str = c("B", "E", "H")
)
# Read in the model object
mod5_fold <- readRDS("../outputs/best_mods_all.rds")$mod5_fold
# Predict and output the result
out <- predict(mod5_fold, newdata = dat, interval = "prediction")
upr_bound <- out[,"upr"]
print(out)
cat("upper bounds: ", upr_bound)
```


## Model 5 for binding energy

Use this model for average ddG from MD + FoldX workflow.
Expected input data are:
- Averaged ddG values for each of these FoldX term, energy_vdw
- Standard deviation of the averaged ddG terms: backbone_vdwclash, energy_vdwclash, entropy_sidec, total
- Secondary structure according to DSSP ("B", "E", "G", "H", "NONE", "S", "T")
- RSA calcuated as described in the manuscript and 01_generate-organized-tables.Rmd, using a method from Tien et al.

```{r}
# Create the input data. Replace the simulated values with your own data
dat <- data.frame(
  energy_vdwclash = rnorm(3, 1, 1),
  backbone_vdwclash_sd = rnorm(3, 0, 1),
  energy_vdwclash_sd = rnorm(3, 0, 1),
  entropy_sidec_sd = rnorm(3, 0, 1),
  total_sd = rnorm(3, 0, 1),
  str = c("B", "E", "H"),
  RSA = c(.2, .2, .8)
)
# Read in the model object
mod5_bind <- readRDS("../outputs/best_mods_all.rds")$mod5_bind
# Predict and output the result
out <- predict(mod5_bind, newdata = dat, interval = "prediction")
upr_bound <- out[,"upr"]
print(out)
cat("upper bounds: ", upr_bound)
```







